---
title: "Bear Den Data Summary"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include = FALSE, warning = FALSE}
library(xaringanthemer)
style_solarized_light(header_color = "#859900")
```


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

# Set working directory to the project folder, so the script can find the targets
#knitr::opts_knit$set(root.dir = rstudioapi::getActiveProject())

# Global knitr options
knitr::opts_chunk$set(warning = FALSE, # don't print code warnings in slides 
                      message = FALSE, # don't print code messages in slides
                      fig.align='center', # center align any figures
                      dev = "svg") # make any plots "svg" format (rather than PNG or JPEG for example)

# Load packages
library(sf)
library(leaflet)
library(dplyr)
library(magrittr)
library(ggplot2)
library(targets)

# Read in tokens
# For this specific markdown file, we need the 'jawg_token' to load
# the nice basemap for the interactive map.
source("../temp/token.R")

# Load targets
# Since this markdown file isn't actually part of the targets
# pipeline, you need this weird workaround to load a target:
# https://github.com/ropensci/targets/issues/230

# dens = dens metadata (with no field visit info attached)
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(dens))

# f = field visits table (with no den metadata attached)
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(f))

# f_full = full field visits table merged with den metadata (incl. spatial lat/lon)
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(f_full))

# v_f = forestry verifications output
v_f <- withr::with_dir(rprojroot::find_root('_targets.R'),
                       tar_read(forestry_verifications_full)) # use `tar_read` when you want to read a target but assign it to a new object name in the R environment

# prop_1km = proportion forested within 1.5km
prop_1km <- withr::with_dir(rprojroot::find_root('_targets.R'),
                            tar_read(prct_age_class_1.5km)) # use `tar_read` when you want to read a target but assign it to a new object name in the R environment

# road_density = road density around each den
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(road_density))

```

```{r dens prep, include = FALSE}
# Wrap this up into `clean_bears` func at some point
# Coerce factors
factors <- c("struct_stage", "age_class", "slope_position", "den_type", "den_tree_species", "district")
dens %<>% mutate_at(factors, factor)
# Remove 999
dens %<>%
  mutate(across(where(is.numeric), ~replace(., . == 999, NA)),
         across(where(is.numeric), ~replace(., . == -999, NA)))

```

There are currently **`r nrow(dens)`** dens being tracked, and a total of **`r nrow(f)`** field visits completed to date (**`r Sys.Date()`**).

```{r den map, out.width='100%', fig.height=6.5, eval=require('leaflet'), echo = FALSE}

dens %>%
  st_transform(4326) %>%
  leaflet() %>% 
  addTiles() %>%
  setView(lng = -127.8, lat = 51.5, zoom = 6) %>%
  addCircleMarkers(layerId = ~den_id,
                   radius = 4,
                   color = "#c852ef",
                   stroke = FALSE,
                   fillOpacity = 0.8,
                   popup = ~htmltools::htmlEscape(den_id)) %>%
  addTiles("Jawg.Terrain", 
           options = providerTileOptions(accessToken = jawg_token),
           group = "Terrain") %>%
  addProviderTiles("Esri.WorldImagery",
                   group = "Satellite") %>%
  addLayersControl(baseGroups = c("Terrain", "Satellite"),
                   position = "topright") %>%
  addMeasure(primaryLengthUnit = "meters")
  
  
```

.footnote[*Use the menus on the top-right to change the basemap or perform measurements.*]

---

# Explore categorical variables
### District

```{r}
knitr::kable(table(dens$district), col.names = c("District", "N"))
```

---

# Explore categorical variables
### Den Tree Species

```{r}
knitr::kable(table(dens$den_tree_species), col.names = c("Den Tree Species", "N"))
```

---

# Explore categorical variables
### Proportion by tree species

```{r}
trees <- table(dens$den_tree_species)
(trees/sum(trees))*100
```

---

# Explore categorical variables
### Den Type

```{r}
knitr::kable(table(dens$den_type), col.names = c("Den Type", "N"))
```

---

# Explore categorical variables
### Age Class

```{r}
knitr::kable(table(dens$age_class), col.names = c("Age Class", "N"))
```

---

# Explore categorical variables
### Den State (initial visit)

```{r}
knitr::kable(table(dens$den_state), col.names = c("Den State", "N"))
```

---
class: center, middle, inverse

# Den Status

## Rates of Re-Use

---

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N"))
```

---
class: center

### "Active"

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N")) |> 
  kableExtra::kable_styling() |> 
  kableExtra::row_spec(c(1:3), bold = TRUE, background = "#b58900") |>
  kableExtra::kable_styling(font_size = 14)
```

---
class: center

### "Non-Active"

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N")) |> 
  kableExtra::kable_styling() |> 
  #kableExtra::row_spec(c(1:4), bold = FALSE, background = "#b58900") |>
  kableExtra::row_spec(c(4:7), bold = TRUE, background = "#93a1a1") |>
  kableExtra::kable_styling(font_size = 14)
```

---
class: center

### "Unknown/Obsolete"

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N")) |> 
  kableExtra::kable_styling() |> 
  #kableExtra::row_spec(c(1:4), bold = FALSE, background = "#b58900") |>
  #kableExtra::row_spec(c(5:14), bold = FALSE, background = "#93a1a1") |>
  kableExtra::row_spec(c(8:9), bold = TRUE, color = "white", background = "#cb4b16") |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

First need to categorize dens into `Active`, `Not Active`, or `Unknown`.
Based on the highlighted rows in the tables of the previous slides, we have:

```{r, echo = FALSE}
# Add a `active_yn` column
f <- f |> dplyr::mutate(active_yn = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                                     den_status == "Obsolete" ~ "Obsolete",
                                                     den_status == "Unknown" ~ "Unknown",
                                                     TRUE ~ "Active"))

# Double check our categories
unique(f[,c("den_status", "active_yn")])
```

---

# Rates of re-use

First need to categorize dens into `Active`, `Not Active`, or `Unknown`.
Based on the highlighted rows in the tables of the previous slides, we have:

```{r}
plyr::count(f$active_yn) |>
  knitr::kable(col.names = c("Den Status", "N"))
```

Total number of visits = **`r nrow(f)`**

---

# Rates of re-use

First need to categorize dens into `Active`, `Not Active`, or `Unknown`.
Based on the highlighted rows in the tables of the previous slides, we have:

```{r}
plyr::count(f$active_yn) |>
  dplyr::mutate(freq = round(freq/nrow(f) * 100)) |>
  knitr::kable(col.names = c("Den Status", "%"))
```

Total number of visits = **`r nrow(f)`**

---

# Rates of re-use

For now, I'm counting `Unknown`s the same as `Not Active`.

This is a quick and dirty assignment of 're-use' - it doesn't take into account if there's been a large temporal gap (e.g., years) between visits.

```{r}
# Arrange table by den_id and date_inspected
f <- f[order(f$den_id, f$date_inspected),]
# Add in a cumulative visit col by each den - i.e. the Nth field visit to the given den
f <- f |>
  dplyr::group_by(den_id) |>    
  dplyr::mutate(cumulative_visit = cumsum(!is.na(den_id)))
# Create `reuse_yn` col
f <- f |> 
  dplyr::group_by(den_id) |>
  dplyr::mutate(reuse_yn = ifelse(cumulative_visit == 1,
                                  # If it's the first visit to the den, set `reuse_yn` == NA
                                  NA, 
                                  # Else if it's NOT the first visit, and the previous record says "Active", set `reuse_yn` == TRUE
                                  (dplyr::lag(active_yn) == "Active" & active_yn == "Active")) 
                )
```

---

#### Rates of re-use

```{r, echo = FALSE}
f[,c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |> 
  DT::datatable(options = list(
    pageLength = 8,
    scrollX = FALSE,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'font-size': '50%'});",
      "}"
      )
    )) |>
  DT::formatStyle(columns = c(1:6), fontSize = "50%") |>
  DT::formatStyle(columns = c(1:6), width = c("5%"))
```

---

# Rates of re-use

#### Count of re-use

.footnote[*Note that `NA` signifies the first visit to a den - we don't know if `reuse_yn == TRUE` or `FALSE` bc it's the first visit*]

```{r}
knitr::kable(plyr::count(f$reuse_yn), col.names = c("Re-Used (T/F)", "N"))
```

---

# Rates of re-use

#### Percent of re-use

.footnote[*Note that `NA` signifies the first visit to a den - we don't know if `reuse_yn == TRUE` or `FALSE` bc it's the first visit*]

```{r}
knitr::kable(plyr::count(f$reuse_yn) |> dplyr::mutate(freq = round(freq/nrow(f) * 100)), col.names = c("Re-Used (T/F)", "%"))
```

---

# Rates of re-use

### ADA_EveRiver_1

```{r, echo = FALSE}
f[f$den_id == "ADA_EveRiver_1", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### COU_CousCreek_2

```{r, echo = FALSE}
f[f$den_id == "COU_CousCreek_2", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### LOW_FlorenceCreek_1

```{r, echo = FALSE}
f[f$den_id == "LOW_FlorenceCreek_1", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---
# Rates of re-use

### TSI_MountRussell_2

```{r, echo = FALSE}
f[f$den_id == "TSI_MountRussell_2", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### NAK_JohnstoneStrait_3

```{r, echo = FALSE}
f[f$den_id == "NAK_JohnstoneStrait_3", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### SAN_PalmerstonRiver_2

```{r, echo = FALSE}
f[f$den_id == "SAN_PalmerstonRiver_2", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "active_yn", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```


---
class: center, middle, inverse

# Forestry summary data

Using the latest verifications (with the caveat that they themselves haven't been fully verified yet)!

---

### Proportion forested within 60m

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(v_f$prop_forest_60m)
```

```{r, fig.height=4}
hist(v_f$prop_forest_60m, main = "Histogram of Proportion Forested", xlab = "Proportion Forested within 60m (%)")
```

---

### Distance to <40 yo forest

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(v_f$dist_lt40)
```

```{r, fig.height=4}
hist(v_f$dist_lt40, main = "Histogram of Distance to <40yo Forest", xlab = "Distance to <40 yo forest (m)")
```

---

### Distance to <40 yo forest

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(v_f$dist_gt40)
```

```{r, fig.height=4}
hist(v_f$dist_gt40, main = "Histogram of Distance to >40yo Forest", xlab = "Distance to >40 yo forest (m)")
```

---

### Distance to nearest road

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(v_f$dist_road)
```

```{r, fig.height=4}
hist(v_f$dist_road, main = "Histogram of Distance to Nearest Road", xlab = "Distance to nearest road (m)")
```


---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r}
# The age class cols are already percentages for each year. So, just
# take the mean percentage of each age class across all years
round(colSums(prop_1km[,4:12]) / nrow(prop_1km)) |>
  knitr::kable(col.names = c("Stand Age Class", "Percentage"))
```
---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r, echo = FALSE}
prop_1km |>  
  # First pivot the data into long format - age class becomes a single column
  tidyr::pivot_longer(cols = 4:12, names_to = "age_class", names_prefix = "age_class_") |>
  # Then summarize the mean % age class by year 
  dplyr::group_by(year, age_class) |> 
  dplyr::summarise(value = mean(value)) |>
  # Then plot
  ggplot(aes(x = year, y = value, fill = age_class)) +
  geom_area() +
  theme_minimal()
```
---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r, echo = FALSE}
prop_1km |>  
  # First pivot the data into long format - age class becomes a single column
  tidyr::pivot_longer(cols = 4:12, names_to = "age_class", names_prefix = "age_class_") |>
  # Condense age class for each den visit into either ">= 3" or "< 3"
  dplyr::mutate(age_class = ifelse(grepl("1|2", age_class), "Less than 3", "Greater or equal to 3")) |>
  dplyr::mutate(age_class = factor(age_class, levels = c("Less than 3", "Greater or equal to 3"))) |> # Flip plotting order so it stays lined up with previous graph
  dplyr::group_by(sample_id, year, age_class) |>
  dplyr::summarise(value = sum(value)) |>
  # Then summarize the mean % age class by year
  dplyr::group_by(year, age_class) |>
  dplyr::summarise(value = mean(value)) |>
  # Then plot
  ggplot(aes(x = year, y = value, fill = age_class)) +
  geom_area() +
  theme_minimal()
```
---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r, echo = FALSE}
prop_1km |>  
  # First pivot the data into long format - age class becomes a single column
  tidyr::pivot_longer(cols = 4:12, names_to = "age_class", names_prefix = "age_class_") |>
  dplyr::mutate(age_class = as.numeric(age_class)) |>
  # Condense age class for each den visit into either ">= 3" or "< 3"
  dplyr::mutate(age_class = dplyr::case_when(age_class < 3 ~ "Less than 3",
                                             age_class >= 8 ~ "Older than 7",
                                             TRUE ~ "3 to 7")) |>
  dplyr::mutate(age_class = factor(age_class, levels = c("Less than 3", "3 to 7", "Older than 7"))) |>
  dplyr::group_by(sample_id, year, age_class) |>
  dplyr::summarise(value = sum(value)) |>
  # Then summarize the mean % age class by year
  dplyr::group_by(year, age_class) |>
  dplyr::summarise(value = mean(value)) |>
  # Then plot
  ggplot(aes(x = year, y = value, fill = age_class)) +
  geom_area() +
  theme_minimal()
```
---

### Road density within 1.5 km

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(road_density$road_density_m2)
```

```{r, fig.height=4}
hist(road_density$road_density_m2, main = "Histogram of Road Density", xlab = "Road Density within 1.5km radius")
```
---

### Road density within 1.5 km

Note this includes all field visits, including dens with repeated visits.

```{r, echo = FALSE}
road_density |>
  ggplot(aes(x = year, y = road_density_m2, color = den_id)) +
  geom_line() +
  geom_point() +
  labs(title = "Road density for each den through time") +
  theme_minimal() +
  theme(legend.position = "none")
```

---
class: center, middle, inverse

# Forestry vs. Den Status

---

```{r boxplot prep, echo = FALSE, dev.args=list(bg="transparent")}
p_f <- merge(v_f, f, by = "sample_id")
#hrbrthemes::import_plex_sans()

ggplot(p_f, aes(x = active_yn, 
                y = prop_forest_60m,
                color = active_yn)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) + 
  labs(x = "Den Status",
       y = "Proportion Forested within 60m (%)")
```


---

```{r, echo = FALSE, dev.args=list(bg="transparent")}
#p_f[p_f$new_dist_lt40 < 2000, ] |> 
ggplot(p_f, aes(x = active_yn, 
                y = dist_lt40,
                color = active_yn)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  scale_y_log10() +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) +
  labs(x = "Den Status",
       y = "LOG Distance to <40yo forest (m)")
```


---

```{r, echo = FALSE, dev.args=list(bg="transparent")}
ggplot(p_f, aes(x = active_yn, 
                y = dist_gt40,
                color = active_yn)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  scale_y_log10() +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) +
  labs(x = "Den Status",
       y = "LOG Distance to >40yo forest (m)")
```


---

```{r, echo = FALSE, dev.args=list(bg="transparent")}
ggplot(p_f, aes(x = active_yn, 
                y = dist_road,
                color = active_yn)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  scale_y_log10() +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) +
  labs(x = "Den Status",
       y = "LOG Distance to nearest road")
```
