---
title: "Bear Den Data Summary"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r xaringan-themer, include = FALSE, warning = FALSE}
library(xaringanthemer)
style_solarized_light(header_color = "#859900")
```


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

# Set working directory to the project folder, so the script can find the targets
#knitr::opts_knit$set(root.dir = rstudioapi::getActiveProject())

# Global knitr options
knitr::opts_chunk$set(warning = FALSE, # don't print code warnings in slides 
                      message = FALSE, # don't print code messages in slides
                      echo = FALSE, # don't print the code itself in any of the slides (unless manually overridden)
                      fig.width = 9, # make default figure width 7"
                      fig.align='center', # center align any figures
                      dev = "svg", # make any plots "svg" format (rather than PNG or JPEG for example)
                      dev.args = list(bg = "transparent") # make the background of all plots transparent
                      ) 

# Load packages
library(targets)  # load up data targets
library(sf)       # spatial manipulation
library(leaflet)  # interactive map
library(dplyr)    # data manipulation
library(magrittr) # data manipulation
library(ggplot2)  # plotting
library(ggsignif) # plotting significance/t-test/wilcox test on boxplots
library(khroma)   # plotting color schemes

# Read in tokens
# For this specific markdown file, we need the 'jawg_token' to load
# the nice basemap for the interactive map.
withr::with_dir(rprojroot::find_root('_targets.R'), source("temp/token.R"))

# Load targets
# Since this markdown file isn't actually part of the targets
# pipeline, you need this weird workaround to load a target:
# https://github.com/ropensci/targets/issues/230

# dens = dens metadata (with no field visit info attached)
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(dens))

# f = field visits table (with no den metadata attached)
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(f))

# f_analysis = field visits data, rearranged such that the current year's
# den status is matched to *the previous year's* forestry data
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(f_analysis))

# prop_1km = proportion forested within 1.5km
prop_1km <- withr::with_dir(rprojroot::find_root('_targets.R'),
                            tar_read(prct_age_class_1.5km)) # use `tar_read` when you want to read a target but assign it to a new object name in the R environment

# road_density = road density around each den
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(road_density))
road_density <- merge(f[3], road_density, all.x = TRUE) # merge in the other sample id's (with zero road density) into road_density dataset
road_density$road_density_m2 <- ifelse(is.na(road_density$road_density_m2), 0, road_density$road_density_m2)

# Define list of "for sure dens"
for_sure_dens <- c("Active in last denning season", 
                   "Not active in last season",
                   "Not active in last season, but recent use (1-4 seasons)",
                   "Not active in last season, no recent use (>4 seasons)",
                   "No recent evidence of use (>4 seasons)")
```

```{r dens prep, include = FALSE}
# Wrap this up into `clean_bears` func at some point
# Coerce factors
factors <- c("struct_stage", "age_class", "slope_position", "den_type", "den_tree_species", "district")
dens %<>% mutate_at(factors, factor)
# Remove 999
dens %<>%
  mutate(across(where(is.numeric), ~replace(., . == 999, NA)),
         across(where(is.numeric), ~replace(., . == -999, NA)))

# Add latitude, longitude cols to data
dens <- st_transform(dens, 3005) # to BC Albers projection
dens <- cbind(dens, st_coordinates(dens))
names(dens)[grep("X", names(dens))] <- "longitude" # technically easting/northing with BC Albers but shhh
names(dens)[grep("Y", names(dens))] <- "latitude"

# Assign either VI or HG to data, to examine by island
dens$region <- ifelse(dens$latitude > 800000, "HG", "VI")
```

```{r f prep, include = FALSE}
# Rename forestry columns to more consistent/easier to use names
f <- dplyr::rename(f,
              "nearest_tree_m" = "distance_nearest_tree_field",
              "f_prop_forest_60m" = "proportion_forested_field",
              "v_prop_forest_60m" = "proportion_forested",
              "f_dist_lt40" = "distance_less40yr_forest_field", 
              "v_dist_lt40" = "v_distance_less40yr_forest",
              "f_dist_gt40" = "distance_grtr40yr_forest_field", 
              "v_dist_gt40" = "v_distance_grtr40year_forest", 
              "f_dist_road" = "distance_nearest_road", 
              "v_dist_road" = "v_distance_nearest_road",
              "windthrow_prct" = "proportion_tree_windthrown", 
              "windthrow_code" = "x_windthrow_code")

# Make just one 'distance from edge' variable
f$dist_from_edge <- ifelse(f$f_dist_lt40 > 0, f$f_dist_lt40, f$f_dist_gt40)
f$log_dist_from_edge <- log(f$dist_from_edge + 1)

# Next, say if it's in old, new, or edge of forest.
f <- f |> dplyr::mutate(f_dist_lt40 = round(f_dist_lt40, 0),
                        f_dist_gt40 = round(f_dist_gt40, 0),
                        age = dplyr::case_when((f_dist_lt40 == f_dist_gt40) ~ "on_edge",
                                         (f_dist_lt40 > f_dist_gt40) ~ "within_old",
                                         (f_dist_lt40 < f_dist_gt40) ~ "within_young"))

# Add a `den_status_binary` column
f <- f |> dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                                     den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                                     den_status == "Obsolete" ~ "Obsolete",
                                                     den_status == "Unknown" ~ "Unknown",
                                                     TRUE ~ "Unknown"))

# Double check our categories
unique(f[,c("den_status", "den_status_binary")]) |> 
  dplyr::arrange(den_status_binary) |>
  knitr::kable()
```


```{r f_analysis prep, include = FALSE}
# Rename forestry columns to more consistent/easier to use names
f_analysis <- dplyr::rename(f_analysis,
              "nearest_tree_m" = "distance_nearest_tree_field",
              "f_prop_forest_60m" = "proportion_forested_field",
              "v_prop_forest_60m" = "proportion_forested",
              "f_dist_lt40" = "distance_less40yr_forest_field", 
              "v_dist_lt40" = "v_distance_less40yr_forest",
              "f_dist_gt40" = "distance_grtr40yr_forest_field", 
              "v_dist_gt40" = "v_distance_grtr40year_forest", 
              "f_dist_road" = "distance_nearest_road", 
              "v_dist_road" = "v_distance_nearest_road",
              "windthrow_prct" = "proportion_tree_windthrown", 
              "windthrow_code" = "x_windthrow_code")

# Next, make just one 'distance from edge' variable
f_analysis$dist_from_edge <- ifelse(f_analysis$f_dist_lt40 > 0, f_analysis$f_dist_lt40, f_analysis$f_dist_gt40)
f_analysis$log_dist_from_edge <- log(f_analysis$dist_from_edge + 1)

# Next, say if it's in old, new, or edge of forest.
f_analysis <- f_analysis |> dplyr::mutate(f_dist_lt40 = round(f_dist_lt40, 0),
                              f_dist_gt40 = round(f_dist_gt40, 0),
                              age = dplyr::case_when((f_dist_lt40 == f_dist_gt40) ~ "on_edge",
                                               (f_dist_lt40 > f_dist_gt40) ~ "within_old",
                                               (f_dist_lt40 < f_dist_gt40) ~ "within_young"))

# Add a `den_status_binary` column
f_analysis <- f_analysis |> dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                                     den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                                     den_status == "Obsolete" ~ "Obsolete",
                                                     den_status == "Unknown" ~ "Unknown",
                                                     TRUE ~ "Unknown"))

# Merge spatial/static den data with analysis subset of field visits
f_analysis <- merge(dens, f_analysis, by = "den_id")

# Merge landscape scale forestry metrics with the f_analysis dataset
# Merge % age class to `f_analysis`
f_analysis <- merge(f_analysis, prop_1km, by.x = "sample_id_forest", by.y = "sample_id")
f_analysis <- dplyr::select(f_analysis, -den_id.y, -year.y)
names(f_analysis)[grep("den_id.x", names(f_analysis))] <- "den_id"
names(f_analysis)[grep("year.x", names(f_analysis))] <- "year"

# Note the all.x = TRUE. If the den isn't in
# the road density df, the road density is 0.
f_analysis <- merge(f_analysis, road_density, 
                    by.x = "sample_id_forest", 
                    by.y = "sample_id", 
                    all.x = TRUE) # again, merging GIS forestry data for the year PRIOR to the den status year
f_analysis <- dplyr::select(f_analysis, -den_id.y, -year.y)
names(f_analysis)[grep("den_id.x", names(f_analysis))] <- "den_id"
names(f_analysis)[grep("year.x", names(f_analysis))] <- "year"

f_analysis$road_density_m2 <- ifelse(is.na(f_analysis$road_density_m2), 0, f_analysis$road_density_m2)

# We'll use the den year sample_id (rather than the forestry year sample id)
# as the default sample id
f_analysis$sample_id <- f_analysis$sample_id_den

# Subset to only columns we care about
f_analysis <- f_analysis |> 
  dplyr::select(sample_id, den_id, den_status, den_status_binary,
                sample_id_forest, sample_id_den,
                date_inspected_den, date_inspected_forest, year,
                region, latitude, longitude, 
                struct_stage, age_class, canopy_closure,
                elevation_m, slope_pct, slope_aspect,
                bed_depth, bed_width, bed_length,
                hair_in_bed,
                forestry_treatment_desc,
                age_class,
                f_prop_forest_60m, f_dist_lt40, f_dist_gt40, f_dist_road,
                windthrow_prct, windthrow_code,
                age_class_1:road_density_m2)
```

```{r f_full prep, include = FALSE}
f_full <- merge(dens, f, by = "den_id")
```

There are currently **`r length(unique(f$den_id))`** dens being tracked, and a total of **`r nrow(f)`** field visits completed to date (**`r Sys.Date()`**).

```{r den map, out.width='100%', fig.height=6.5, eval=require('leaflet'), echo = FALSE}
dens %>%
  st_transform(4326) %>%
  leaflet() %>% 
  addTiles() %>%
  setView(lng = -127.8, lat = 51.5, zoom = 6) %>%
  addCircleMarkers(layerId = ~den_id,
                   radius = 4,
                   color = "#c852ef",
                   stroke = FALSE,
                   fillOpacity = 0.8,
                   popup = ~htmltools::htmlEscape(den_id)) %>%
  addTiles("Jawg.Terrain", 
           options = providerTileOptions(accessToken = jawg_token),
           group = "Terrain") %>%
  addProviderTiles("Esri.WorldImagery",
                   group = "Satellite") %>%
  addLayersControl(baseGroups = c("Terrain", "Satellite"),
                   position = "topright") %>%
  addMeasure(primaryLengthUnit = "meters")
  
  
```

.footnote[*Use the menus on the top-right to change the basemap or perform measurements.*]

---

# Overall summary

```{r, dev.args=list(bg="transparent")}
f |> 
  dplyr::mutate(year = lubridate::year(date_inspected)) |>
  ggplot(aes(x = year)) +
  geom_bar() +
  scale_x_continuous(breaks = 2014:2024) +
  labs(x = "Year",
       y = "N field visits",
       title = "Number of field visits per year") +
  theme_minimal() 
```

---

# Overall summary

```{r, dev.args=list(bg="transparent")}
f |> 
  dplyr::mutate(year = lubridate::year(date_inspected)) |>
  dplyr::filter(grepl("Yes", monitored_by_camera)) |>
  ggplot(aes(x = year)) +
  geom_bar() +
  scale_x_continuous(limits = c(2014,2024.5), 
                     breaks = 2014:2024) +
  labs(x = "Year",
       y = "N deployed cameras",
       title = "Number of deployed cameras per year") +
  theme_minimal() 
```


---

# Overall summary

### Den Status - what have we got?

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N"))
```

---

# Overall summary

```{r, dev.args=list(bg="transparent")}
f |> 
  dplyr::filter(den_status %in% c("Currently active", "Active in last denning season")) |>
  dplyr::mutate(year = lubridate::year(date_inspected)) |>
  ggplot(aes(x = year)) +
  geom_bar() +
  scale_x_continuous(breaks = 2014:2024) +
  labs(x = "Year",
       y = "N active dens",
       title = "Number of dens determined to be 'Active' per year") +
  theme_minimal() 
```

---

# Where are our dens?
### District

```{r, echo = FALSE}
knitr::kable(table(dens$district), col.names = c("District", "N"))
```

---

# Where are our dens?
### Region

```{r, fig.height = 5.5, dev.args=list(bg="transparent")}
dens |> 
  ggplot(aes(x = region)) +
  geom_bar() +
  labs(x = "Region",
       y = "N dens",
       title = "Number of dens in each region") +
  theme_minimal() 
```

---

# Where are our dens?
### District

```{r, fig.height = 5.5, dev.args=list(bg="transparent")}
dens |> 
  ggplot(aes(x = district)) +
  geom_bar() +
  labs(x = "District",
       y = "N dens",
       title = "Number of dens in each district") +
  theme_minimal() 
```

---
class: center, middle, inverse

# What does a den look like?

---

# What does a den look like?
### Den Tree Species

```{r, echo = FALSE}
knitr::kable(table(dens$den_tree_species), col.names = c("Den Tree Species", "N"))
```

---

# What does a den look like?
### Proportion by tree species

```{r, echo = FALSE}
trees <- table(dens$den_tree_species)
knitr::kable(round((trees/sum(trees))*100, 0), col.names = c("Den Tree Species", "%"))
```

---

# What does a den look like?
### Den Type

```{r, echo = FALSE}
knitr::kable(table(dens$den_type), col.names = c("Den Type", "N"))
```

---

# What does a den look like?
### DBH

```{r, fig.height = 5.5}
hist(dens$den_tree_dbh, main = "Histogram of Den Tree DBH", xlab = "DBH (cm)")
```

---

# What does a den look like?
### Chamber width, height, and length

```{r, fig.height = 5.5, dev.args=list(bg="transparent")}
dens |>
  sf::st_drop_geometry() |>
  dplyr::select(chamber_width, chamber_height, chamber_length) |>
  tidyr::pivot_longer(cols = 1:3,
                      names_to = "Measurement",
                      values_to = "cm") |>
  ggplot(aes(x = cm)) +
  geom_density(aes(color = Measurement,
                   fill = Measurement),
               alpha = 0.1) +
  scale_color_manual(values = c("#E69F00", "#009E73", "#CC79A7"),
                     name = "Measurement",
                     labels = c("Chamber height",
                                "Chamber length",
                                "Chamber width")) +
  scale_fill_manual(values = c("#E69F00", "#009E73", "#CC79A7"),
                    name = "Measurement",
                    labels = c("Chamber height",
                               "Chamber length",
                               "Chamber width")) +
  labs(x = "Measurement (cm)",
       y = "Density") +
  theme_minimal()
```

---

# What does a den look like?
### Dens with Bed Cups during each field visit (N)

```{r}
knitr::kable(table(f$bedding_cup_present), col.names = c("Bedding Cup Present?", "N"))
```

---

# What does a den look like?
### Dens with Bed Cups during each field visit (%)

```{r}
cups <- table(f$bedding_cup_present)
knitr::kable(round(cups/sum(cups)*100, 0), col.names = c("Bedding Cup Present?", "%"))
```

---

# What does a den look like?
### Hair in Den during each field visit (N)

```{r}
knitr::kable(table(f$hair_in_bed), col.names = c("Hair in Den?", "N"))
```

---

# What does a den look like?
### Hair in Den during each field visit (%)

```{r}
hair <- table(f$hair_in_bed)
knitr::kable(round(hair/sum(hair)*100, 0), col.names = c("Hair in Den?", "%"))
```

---

# What does a den look like?
### Age Class

```{r, echo = FALSE}
knitr::kable(table(dens$age_class), col.names = c("Age Class", "N"))
```

---

# What does a den look like?
### Den State (initial visit)

```{r, echo = FALSE}
knitr::kable(table(dens$den_state), col.names = c("Den State", "N"))
```

--- 

# What's the geography like?
### Elevation

```{r}
hist(dens$elevation_m, main = "Histogram of den elevations", xlab = "Elevation (m)")
```

--- 

# What's the geography like?
### Slope grade (%)

```{r}
hist(dens$slope_pct, main = "Histogram of slope grades", xlab = "Slope (%)")
```

--- 

# What's the geography like?
### Slope aspect

```{r}
hist(dens$slope_aspect, main = "Histogram of slope aspects", xlab = "Slope aspect (degrees)",
     xlim = range(0, 360), breaks = seq(0, 360, 45))
```

---
class: center, middle, inverse

# Den Status

## Rates of Re-Use

---

### Den Status - what have we got?

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N"))
```

---
class: center

### "Active"

Dens with the highlighted status were considered "confirmed active" within the last season.

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N")) |> 
  kableExtra::kable_styling() |> 
  kableExtra::row_spec(c(1, 3), bold = TRUE, background = "#b58900") |>
  kableExtra::kable_styling(font_size = 14)
```

---
class: center

### "Non-Active"

Dens with the highlighted status were considered "confirmed *not* active" within the last season.

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N")) |> 
  kableExtra::kable_styling() |> 
  #kableExtra::row_spec(c(1:4), bold = FALSE, background = "#b58900") |>
  kableExtra::row_spec(c(4:7), bold = TRUE, background = "#93a1a1") |>
  kableExtra::kable_styling(font_size = 14)
```

---
class: center

### "Unknown/Obsolete"

Dens with the highlighted status were excluded from den status assessments.

```{r, echo=FALSE}
knitr::kable(table(f$den_status), col.names = c("Den Status", "N")) |> 
  kableExtra::kable_styling() |> 
  #kableExtra::row_spec(c(1:4), bold = FALSE, background = "#b58900") |>
  #kableExtra::row_spec(c(5:14), bold = FALSE, background = "#93a1a1") |>
  kableExtra::row_spec(c(2, 8:9), bold = TRUE, color = "white", background = "#cb4b16") |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

First need to categorize dens into `Active`, `Not Active`, or `Unknown`.
Based on the highlighted rows in the tables of the previous slides, we have:

```{r den_status_binary, echo = FALSE}
# Add a `den_status_binary` column
f <- f |> dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                                     den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                                     den_status == "Obsolete" ~ "Obsolete",
                                                     den_status == "Unknown" ~ "Unknown",
                                                     TRUE ~ "Unknown"))

# Double check our categories
unique(f[,c("den_status", "den_status_binary")]) |> 
  dplyr::arrange(den_status_binary) |>
  knitr::kable()
```


---

# Rates of re-use

First need to categorize dens into `Active`, `Not Active`, or `Unknown`.
Based on the highlighted rows in the tables of the previous slides, we have:

```{r}
plyr::count(f$den_status_binary) |>
  knitr::kable(col.names = c("Den Status", "N"))
```

Total number of visits = **`r nrow(f)`**

---

# Rates of re-use

First need to categorize dens into `Active`, `Not Active`, or `Unknown`.
Based on the highlighted rows in the tables of the previous slides, we have:

```{r}
plyr::count(f$den_status_binary) |>
  dplyr::mutate(freq = round(freq/nrow(f) * 100)) |>
  knitr::kable(col.names = c("Den Status", "%"))
```

Total number of visits = **`r nrow(f)`**

---

# Rates of re-use

For now, I'm counting `Unknown`s the same as `Not Active`.

This is a quick and dirty assignment of 're-use' - it doesn't take into account if there's been a large temporal gap (e.g., years) between visits.

But, essentially, if the status was `Active` for two records in a row, it's counted as *re-used*.

```{r}
# Arrange table by den_id and date_inspected
f <- f[order(f$den_id, f$date_inspected),]
# Add in a cumulative visit col by each den - i.e. the Nth field visit to the given den
f <- f |>
  dplyr::group_by(den_id) |>    
  dplyr::mutate(cumulative_visit = cumsum(!is.na(den_id)))
# Create `reuse_yn` col
f <- f |> 
  dplyr::group_by(den_id) |>
  dplyr::mutate(reuse_yn = ifelse(cumulative_visit == 1,
                                  # If it's the first visit to the den, set `reuse_yn` == NA
                                  NA, 
                                  # Else if it's NOT the first visit, and the previous record says "Active", set `reuse_yn` == TRUE
                                  (dplyr::lag(den_status_binary) == "Active" & den_status_binary == "Active")) 
                )
```

---

#### Rates of re-use

```{r, echo = FALSE}
f[,c("den_id", "sample_id", "date_inspected", "cumulative_visit", "den_status_binary", "reuse_yn")] |> 
  DT::datatable(options = list(
    pageLength = 8,
    scrollX = FALSE,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'font-size': '50%'});",
      "}"
      )
    )) |>
  DT::formatStyle(columns = c(1:6), fontSize = "50%") |>
  DT::formatStyle(columns = c(1:6), width = c("5%"))
```

---

# Rates of re-use

#### Count of re-use

.footnote[*Note that `NA` signifies the first visit to a den - we don't know if `reuse_yn == TRUE` or `FALSE` bc it's the first visit*]

```{r}
knitr::kable(plyr::count(f$reuse_yn), col.names = c("Re-Used (T/F)", "N"))
```

---

# Rates of re-use

#### Percent of re-use

.footnote[*Note that `NA` signifies the first visit to a den - we don't know if `reuse_yn == TRUE` or `FALSE` bc it's the first visit*]

```{r}
knitr::kable(plyr::count(f$reuse_yn) |> dplyr::mutate(freq = round(freq/nrow(f) * 100)), col.names = c("Re-Used (T/F)", "%"))
```

---

# Rates of re-use

### A few example dens...

---

# Rates of re-use

### ADA_EveRiver_1

Was never active in the first place.

```{r, echo = FALSE}
f[f$den_id == "ADA_EveRiver_1", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "den_status_binary", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### COU_CousCreek_2

Active more than once, but not two years in a row. 

```{r, echo = FALSE}
f[f$den_id == "COU_CousCreek_2", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "den_status_binary", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### TSI_MountRussell_2

Was active, but then turned `Obsolete`.

```{r, echo = FALSE}
f[f$den_id == "TSI_MountRussell_2", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "den_status_binary", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```

---

# Rates of re-use

### SAN_RonningCreek_2

Active for two years in a row!

```{r, echo = FALSE}
f[f$den_id == "SAN_RonningCreek_2", c("den_id", "sample_id", "date_inspected", "cumulative_visit", "den_status_binary", "reuse_yn")] |>
  knitr::kable() |>
  kableExtra::kable_styling(font_size = 14)
```


---
class: center, middle, inverse

# Den type/size vs Status

---

### Chamber dimensions vs Den Status

```{r, dev.args=list(bg="transparent")}
label_names <- c(
  `chamber_width` = "Chamber width",
  `chamber_height` = "Chamber height",
  `chamber_length` = "Chamber length"
)

f |>
  dplyr::left_join(dens, by = "den_id") |>
  dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                             den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                             den_status == "Obsolete" ~ "Obsolete",
                                             den_status == "Unknown" ~ "Unknown",
                                             TRUE ~ "Unknown")) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::select(den_status_binary, chamber_width, chamber_height, chamber_length) |>
  tidyr::pivot_longer(cols = c("chamber_width", "chamber_height", "chamber_length"),
                      names_to = "Measurement",
                      values_to = "cm") |>
  ggplot(aes(x = den_status_binary,
             y = cm,
             color = den_status_binary)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) + 
  labs(x = "Den Status",
       y = "Measurement (cm)") +
  facet_wrap(~ Measurement, labeller = as_labeller(label_names))
```

---

### Den type vs Den Status

Root bole type dens disproportionately have more activity.

```{r}
chisq <- f |>
  dplyr::left_join(dens, by = "den_id") |>
  dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                             den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                             den_status == "Obsolete" ~ "Obsolete",
                                             den_status == "Unknown" ~ "Unknown",
                                             TRUE ~ "Unknown")) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::group_by(den_type, den_status_binary) |>
  dplyr::tally() |>
  tidyr::pivot_wider(names_from = den_type,
                     values_from = n) |>
  dplyr::mutate(den_status_binary = dplyr::if_else(!is.na(den_status_binary), den_status_binary, "NA")) |>
  tibble::column_to_rownames(var = "den_status_binary") |>
  dplyr::mutate_all(~replace(., is.na(.), 0)) |>
  as.matrix() |>
  chisq.test()

corrplot::corrplot(chisq$residuals, is.corr = FALSE)
```

---

### Bed Cup vs Den Status

Active dens dispoportionately have a bedding cup (makes sense - it's one of the decision criteria).

```{r}
chisq <- f |>
  dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                             den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                             den_status == "Obsolete" ~ "Obsolete",
                                             den_status == "Unknown" ~ "Unknown",
                                             TRUE ~ "Unknown")) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::group_by(bedding_cup_present, den_status_binary) |>
  dplyr::tally() |>
  tidyr::pivot_wider(names_from = bedding_cup_present,
                     values_from = n) |>
  dplyr::mutate(den_status_binary = dplyr::if_else(!is.na(den_status_binary), den_status_binary, "NA")) |>
  tibble::column_to_rownames(var = "den_status_binary") |>
  dplyr::mutate_all(~replace(., is.na(.), 0)) |>
  as.matrix() |>
  chisq.test()

corrplot::corrplot(chisq$residuals, is.corr = FALSE)
```

---

### Hair in Bed vs Den Status

Active dens dispoportionately have hair in the bed (makes sense - it's one of the decision criteria).

```{r}
chisq <- f |>
  dplyr::mutate(den_status_binary = dplyr::case_when(grepl("No", den_status) ~ "Not Active",
                                             den_status %in% c("Active in last denning season", "Currently active") ~ "Active",
                                             den_status == "Obsolete" ~ "Obsolete",
                                             den_status == "Unknown" ~ "Unknown",
                                             TRUE ~ "Unknown")) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::group_by(hair_in_bed, den_status_binary) |>
  dplyr::tally() |>
  tidyr::pivot_wider(names_from = hair_in_bed,
                     values_from = n) |>
  dplyr::mutate(den_status_binary = dplyr::if_else(!is.na(den_status_binary), den_status_binary, "NA")) |>
  tibble::column_to_rownames(var = "den_status_binary") |>
  dplyr::mutate_all(~replace(., is.na(.), 0)) |>
  as.matrix() |>
  chisq.test()

corrplot::corrplot(chisq$residuals, is.corr = FALSE)
```

---
class: center, middle, inverse

# Geography vs Status

---

### Elevation

```{r}
f_full |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  ggplot(aes(x = den_status_binary, 
             y = elevation_m)) +
  geom_boxplot() +
  geom_jitter() +
  geom_signif(comparisons = list(c("Active", "Not Active")), map_signif_level = TRUE) +
  theme_minimal() +
  labs(x = "Den Status",
       y = "Elevation (m)",
       title = "Den Status vs. Elevation (m)",
       subtitle = "All categories binned into either 'Active' or 'Not active'") 
```

---

### Slope grade (%)

```{r}
f_full |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  ggplot(aes(x = den_status_binary, 
             y = slope_pct)) +
  geom_boxplot() +
  geom_jitter() +
  geom_signif(comparisons = list(c("Active", "Not Active")), map_signif_level = TRUE) +
  theme_minimal() +
  labs(x = "Den Status",
       y = "Slope Grade (%)",
       title = "Den Status vs. Slope (%)",
       subtitle = "All categories binned into either 'Active' or 'Not active'") 

```

---

### Slope aspect (degrees)

```{r}
f_full |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  ggplot(aes(x = den_status_binary, 
             y = slope_aspect)) +
  geom_boxplot() +
  geom_jitter() +
  geom_signif(comparisons = list(c("Active", "Not Active")), map_signif_level = TRUE) +
  theme_minimal() +
  labs(x = "Den Status",
       y = "Slope Aspect",
       title = "Den Status vs. Slope Aspect",
       subtitle = "All categories binned into either 'Active' or 'Not active'") 

```


---
class: center, middle, inverse

# Forestry summary data

---

### Canopy closure

Note this value was not updated past the first field visit.

```{r}
hist(dens$canopy_closure, main = "Histogram of canopy closure", xlab = "Canopy closure (%)", xlim = range(0, 100))
```

---

### Proportion forested within 60m

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(f$f_prop_forest_60m)
```

```{r, fig.height=4}
hist(f$f_prop_forest_60m, main = "Histogram of Proportion Forested", xlab = "Proportion Forested within 60m (%)")
```

---

### Distance to <40 yo forest

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(f$f_dist_lt40)
```

```{r, fig.height=4}
hist(f$f_dist_lt40, main = "Histogram of Distance to <40yo Forest", xlab = "Distance to <40 yo forest (m)")
```

---

### Distance to <40 yo forest

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(f$f_dist_gt40)
```

```{r, fig.height=4}
hist(f$f_dist_gt40, main = "Histogram of Distance to >40yo Forest", xlab = "Distance to >40 yo forest (m)")
```

---

### Distance to nearest road

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(f$f_dist_road)
```

```{r, fig.height=4}
hist(f$f_dist_road, main = "Histogram of Distance to Nearest Road", xlab = "Distance to nearest road (m)")
```


---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r}
# The age class cols are already percentages for each year. So, just
# take the mean percentage of each age class across all years
round(colSums(prop_1km[,4:12]) / nrow(prop_1km)) |>
  knitr::kable(col.names = c("Stand Age Class", "Percentage"))
```

---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r, echo = FALSE, fig.width = 10, dev.args=list(bg="transparent")}
prop_1km |>  
  # First pivot the data into long format - age class becomes a single column
  tidyr::pivot_longer(cols = 4:12, names_to = "age_class", names_prefix = "age_class_") |>
  dplyr::mutate(age_class = factor(age_class, levels = c(1:9))) |>
  # Then summarize the mean % age class by year 
  dplyr::group_by(year, age_class) |> 
  dplyr::summarise(value = mean(value)) |>
  # Then plot
  ggplot(aes(x = year, y = value, fill = age_class)) +
  geom_area() +
  scale_fill_light(reverse = TRUE) +
  theme_minimal()
```

---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r, echo = FALSE, fig.width = 10, dev.args=list(bg="transparent")}
prop_1km |>  
  # First pivot the data into long format - age class becomes a single column
  tidyr::pivot_longer(cols = 4:12, names_to = "age_class", names_prefix = "age_class_") |>
  # Condense age class for each den visit into either ">= 3" or "< 3"
  dplyr::mutate(age_class = ifelse(grepl("1|2", age_class), "Less than 3", "Greater or equal to 3")) |>
  dplyr::mutate(age_class = factor(age_class, levels = c("Less than 3", "Greater or equal to 3"))) |> # Flip plotting order so it stays lined up with previous graph
  dplyr::group_by(sample_id, year, age_class) |>
  dplyr::summarise(value = sum(value)) |>
  # Then summarize the mean % age class by year
  dplyr::group_by(year, age_class) |>
  dplyr::summarise(value = mean(value)) |>
  # Then plot
  ggplot(aes(x = year, y = value, fill = age_class)) +
  geom_area() +
  scale_fill_light(reverse = TRUE) +
  theme_minimal()
```

---

### Proportion forested within 1.5 km

Note this includes all field visits, including dens with repeated visits

```{r, echo = FALSE, fig.width = 10, dev.args=list(bg="transparent")}
prop_1km |>  
  # First pivot the data into long format - age class becomes a single column
  tidyr::pivot_longer(cols = 4:12, names_to = "age_class", names_prefix = "age_class_") |>
  dplyr::mutate(age_class = as.numeric(age_class)) |>
  # Condense age class for each den visit into either ">= 3" or "< 3"
  dplyr::mutate(age_class = dplyr::case_when(age_class < 3 ~ "Less than 3",
                                             age_class >= 8 ~ "Older than 7",
                                             TRUE ~ "3 to 7")) |>
  dplyr::mutate(age_class = factor(age_class, levels = c("Less than 3", "3 to 7", "Older than 7"))) |>
  dplyr::group_by(sample_id, year, age_class) |>
  dplyr::summarise(value = sum(value)) |>
  # Then summarize the mean % age class by year
  dplyr::group_by(year, age_class) |>
  dplyr::summarise(value = mean(value)) |>
  # Then plot
  ggplot(aes(x = year, y = value, fill = age_class)) +
  geom_area() +
  scale_fill_light(reverse = TRUE) +
  theme_minimal()
```

---

### Road density within 1.5 km

Note this includes all field visits, including dens with repeated visits.

```{r}
summary(road_density$road_density_m2)
```

```{r, fig.height=4}
hist(road_density$road_density_m2, main = "Histogram of Road Density", xlab = "Road Density within 1.5km radius")
```

---

### Road density within 1.5 km

It doesn't change much through time... (this will be a theme)

```{r, echo = FALSE, fig.width = 7, dev.args=list(bg="transparent")}
road_density |>
  ggplot(aes(x = year, y = road_density_m2, color = den_id)) +
  geom_line() +
  geom_point() +
  labs(title = "Road density for each den through time") +
  theme_minimal() +
  theme(legend.position = "none")
```

---
class: center, middle, inverse

# Forestry vs. Den Status

---

# Forestry vs. Den Status

Now, we need to do a bit of data wrangling. In theory, the den status of **the current year** is influenced by the forestry data **from the year before**. So, we match the den status with last year's forestry data from our field visits. In doing so, this cuts our dataset own to `r nrow(f_analysis)` records. 

With the exception of canopy closure, all subsequent plots show den status of year *N* versus forestry data from year *N - 1*. Canopy closure was only measured once in the study (at the first den visit) so take that result with a grain of salt.

---

## Den Status vs Canopy Closure

```{r fig.height = 5.5, echo = FALSE, dev.args=list(bg="transparent")}
#hrbrthemes::import_plex_sans()

ggplot(f_full, aes(x = den_status_binary,
                       y = canopy_closure,
                       color = den_status_binary)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) + 
  labs(x = "Den Status",
       y = "Canopy Closure (%)")
```

---

## Den Status vs Proportion Forested within 60m

```{r fig.height = 5.5, echo = FALSE, dev.args=list(bg="transparent")}
#hrbrthemes::import_plex_sans()

ggplot(f_analysis, aes(x = den_status_binary,
                       y = f_prop_forest_60m,
                       color = den_status_binary)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) + 
  labs(x = "Den Status",
       y = "Proportion Forested within 60m (%)")
```


---

## Den Status vs Distance to <40 yo Forest

```{r, echo = FALSE, dev.args=list(bg="transparent")}
ggplot(f_analysis, aes(x = den_status_binary, 
                       y = f_dist_lt40,
                       color = den_status_binary)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  scale_y_log10() +
  annotation_logticks(side = "l") +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) +
  labs(x = "Den Status",
       y = "LOG Distance to <40 yo forest (m)")
```


---

## Den Status vs Distance to >40 yo Forest

```{r, echo = FALSE, dev.args=list(bg="transparent")}
ggplot(f_analysis, aes(x = den_status_binary, 
                       y = f_dist_gt40, 
                       color = den_status_binary)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  scale_y_log10() +
  annotation_logticks(side = "l") +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) +
  labs(x = "Den Status",
       y = "LOG Distance to >40 yo forest (m)")
```


---

## Den Status vs Distance to Nearest Road

```{r, echo = FALSE, dev.args=list(bg="transparent")}
ggplot(f_analysis, aes(x = den_status_binary, 
                       y = f_dist_road,
                       color = den_status_binary)) +
  geom_boxplot(fill = NA) +
  geom_jitter() +
  ggsignif::geom_signif(comparisons = list(c("Active", "Not Active")),
                        map_signif_level = TRUE) +
  scale_y_log10() +
  annotation_logticks(side = "l") +
  #hrbrthemes::theme_ipsum_ps(axis_title_just = "mc") +
  hrbrthemes::scale_color_ipsum() +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent", color = NA_character_),
        plot.background = element_rect(fill = "transparent", color = NA_character_),
        text = element_text(size = 20)) +
  labs(x = "Den Status",
       y = "LOG Distance to nearest road")
```

---
class: center, middle, inverse

# How many dens fall within best practice guidelines?

---

### Dens >75m from the nearest road

```{r}
n_dens <- length(unique(f$den_id))
  
f |>
  dplyr::select(den_id, f_dist_road) |>
  dplyr::filter(!is.na(f_dist_road)) |>
  dplyr::distinct() |>
  dplyr::mutate(over_75 = f_dist_road > 75) |>
  #dplyr::group_by(over_75) |>
  #dplyr::summarize(n_over_75 = dplyr::n()) |>
  ggplot(aes(x = over_75)) +
  geom_bar() +
  labs(title = "N dens >75m from a road",
          x = "Is the den >75m from a road?",
          y = "N dens",
       caption = paste(n_dens, "dens total")) +
  theme_minimal()
```

---

### Dens >75m from the nearest road

```{r}
f |>
  dplyr::select(den_id, den_status_binary, f_dist_road) |>
  dplyr::filter(!is.na(f_dist_road)) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::distinct() |>
  dplyr::mutate(over_75 = f_dist_road > 75) |>
  #dplyr::group_by(over_75) |>
  #dplyr::summarize(n_over_75 = dplyr::n()) |>
  ggplot(aes(x = over_75,
             fill = den_status_binary)) +
  geom_bar() +
  scale_fill_bright(name = "Den status") +
  labs(title = "N dens >75m from a road",
       x = "Is the den >75m from a road?",
       y = "N dens",
       caption = paste(n_dens, "dens total")) +
  facet_wrap(~ den_status_binary) +
  theme_minimal()
```

---

### Dens >30m from cutblock/new growth

```{r}
f |>
  dplyr::select(den_id, f_dist_lt40) |>
  dplyr::filter(!is.na(f_dist_lt40)) |>
  dplyr::distinct() |>
  dplyr::mutate(over_30 = f_dist_lt40 > 30) |>
  ggplot(aes(x = over_30)) +
  geom_bar() +
  labs(title = "N dens >30m from a cutblock",
       x = "Is the den >30m from a cutblock (forest <40 years old)?",
       y = "N dens",
       caption = paste(n_dens, "dens total")) +
  theme_minimal()
```

---

### Dens >30m from cutblock/new growth

```{r}
f |>
  dplyr::select(den_id, den_status_binary, f_dist_lt40) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::filter(!is.na(f_dist_lt40)) |>
  dplyr::distinct() |>
  dplyr::mutate(over_30 = f_dist_lt40 > 30) |>
  ggplot(aes(x = over_30,
             fill = den_status_binary)) +
  geom_bar() +
  scale_fill_bright(name = "Den status") +
  facet_wrap(~ den_status_binary) +
  labs(title = "N dens >30m from a cutblock",
       x = "Is the den >30m from a cutblock (forest <40 years old)?",
       y = "N dens",
       caption = paste(n_dens, "dens total")) +
  theme_minimal()
```

---

### Dens with 100% forest retention in a 60m radius

```{r}
f |>
  dplyr::select(den_id, f_prop_forest_60m) |>
  dplyr::filter(!is.na(f_prop_forest_60m)) |>
  dplyr::distinct() |>
  dplyr::mutate(allforest = f_prop_forest_60m == 100) |>
  ggplot(aes(x = allforest)) +
  geom_bar() +
  labs(title = "N dens 100% forested within 60m",
       x = "Is the den 100% forested in a 60m radius?",
       y = "N dens",
       caption = paste(n_dens, "dens total")) +
  theme_minimal()
```

---

### Dens with 100% forest retention in a 60m radius

```{r}
f |>
  dplyr::select(den_id, den_status_binary, f_prop_forest_60m) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::filter(!is.na(f_prop_forest_60m)) |>
  dplyr::distinct() |>
  dplyr::mutate(allforest = f_prop_forest_60m == 100) |>
  ggplot(aes(x = allforest,
             fill = den_status_binary)) +
  geom_bar() +
  scale_fill_bright(name = "Den status") +
  facet_wrap(~ den_status_binary) +
  labs(title = "N dens 100% forested within 60m",
       x = "Is the den 100% forested in a 60m radius?",
       y = "N dens",
       caption = paste(n_dens, "dens total")) +
  theme_minimal()
```

---

### Dens within each patch size class

```{r}
n_dens <- length(unique(f[["den_id"]][!is.na(f$forestry_treatment_desc)]))
f |>
  dplyr::select(den_id, den_status_binary, forestry_treatment_desc) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::filter(!is.na(forestry_treatment_desc)) |>
  dplyr::distinct() |>
  ggplot(aes(x = forestry_treatment_desc)) +
  geom_bar() +
  scale_fill_bright(name = "Den status") +
  labs(title = "N dens by patch size",
       x = "Patch status",
       y = "N dens",
       caption = paste(n_dens, "dens total with enough information to determine patch size")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip()
```

---

### Dens within each patch size class

```{r}
f |>
  dplyr::select(den_id, den_status_binary, forestry_treatment_desc) |>
  dplyr::filter(den_status_binary %in% c("Active", "Not Active")) |>
  dplyr::filter(!is.na(forestry_treatment_desc)) |>
  dplyr::distinct() |>
  ggplot(aes(x = forestry_treatment_desc,
             fill = factor(den_status_binary))) +
  geom_bar(position = "dodge") +
  scale_fill_bright(name = "Den status") +
  labs(title = "N dens by patch size",
       x = "Patch status",
       y = "N dens",
       caption = paste(n_dens, "dens total with enough information to determine patch size")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip()
```

---
class: center, middle, inverse

# How well are we sampling the landscape?

---

# How well are we sampling the landscape?

The next set of slides will compare points randomly sampled across the landscape versus our dens. Compared to just randomly dropping pins on a map of Vancouver Island and Haida Gwaii, how well are we sampling the landscape with our sample of dens?

```{r pseudo dens setup, include = FALSE}
# pseudo_dens = randomly sampled points on the landscape
# to compare to our actual sampled dens
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(pseudo_dens))

# First, for the pseudo den data, we're going to make
# the same assumption as for the real data - this year's
# den status needs to be paired with last year's forestry.
# Add 1 from the pseudo data `year` so that the 
# data runs from 2019-2023, rather than 2020-2024, 
# because then it will line up with our real data (which
# only runs until 2023.)
pseudo_dens$year <- pseudo_dens$year + 1

# Add latitude, longitude cols to data
pseudo_dens <- cbind(pseudo_dens, st_coordinates(pseudo_dens))
names(pseudo_dens)[grep("X", names(pseudo_dens))] <- "longitude"
names(pseudo_dens)[grep("Y", names(pseudo_dens))] <- "latitude"

# Assign either VI or HG to data, to examine by island
pseudo_dens$region <- ifelse(pseudo_dens$latitude > 800000, "HG", "VI")

##### Assign random den status #####

# Randomly assign a den_status to the pseudo_data, following
# similar proportions as the actual data

# Pull the % distribution of den statuses from the actual
# data, then round to the nearest hundreth. We'll use those
# percentages to simulate den status in the pseudo data.
status_freq <- round(prop.table(table(f_analysis$den_status)), 2)
status_freq

# Generate random den status for our pseudo dens
pseudo_dens$den_status <- sample(names(status_freq), 
                                 nrow(pseudo_dens), 
                                 replace = TRUE, 
                                 prob = status_freq)

# Check that frequencies line up
prop.table(table(pseudo_dens$den_status))

# Make den status a factor
pseudo_dens$den_status <- factor(x = pseudo_dens$den_status,
                                 levels = c("Currently active", #x
                                            "Active in last denning season", #x
                                            "Active recently (0-4 seasons)", #x
                                            "Not active in last season, but recent use (1-4 seasons)", #x
                                            "Not active in last season, no recent use (>4 seasons)", #x
                                            "No recent evidence of use (>4 seasons)", #x
                                            "Not active in last season", #x
                                            "Obsolete", #x
                                            "Unknown")) #x

# Bin into binary categories
pseudo_dens <- pseudo_dens |> 
  dplyr::mutate(den_status_binary = dplyr::case_when(grepl("^No", den_status) ~ "Not active",
                                                     den_status == "Obsolete" ~ "Obsolete",
                                                     den_status == "Unknown" ~ "Unknown",
                                                     TRUE ~ "Active"))

# Pull our parks shapefiles
parks1 <- withr::with_dir(rprojroot::find_root('_targets.R'),
                            tar_read(hg_vi_tantalis_parks))
parks2 <- withr::with_dir(rprojroot::find_root('_targets.R'),
                            tar_read(hg_vi_federal_parks))

names(parks2)[2] <- "PROTECTED_LANDS_NAME"
parks2$PROTECTED_LANDS_DESIGNATION <- "FEDERAL PARK"

parks <- dplyr::bind_rows(parks1, parks2)
parks <- st_as_sf(parks, wkt = "WKT_GEOM")
st_crs(parks) <- 3005
names(parks)[5] <- "geometry" # rename geometry column
st_geometry(parks) <- "geometry"
names(parks) <- tolower(names(parks))

rm(parks1, parks2)

# Intersect parks with dens/pseudo_dens, so we know how
# many pseudo dens were generated inside park lands
pseudo_dens$in_park <- FALSE
pseudo_dens[["in_park"]][unlist(st_intersects(parks, pseudo_dens))] <- TRUE

dens$in_park <- FALSE
dens[["in_park"]][unlist(st_intersects(parks, dens))] <- TRUE

# Merge dens$in_park to f_analysis
f_analysis <- merge(f_analysis, sf::st_drop_geometry(dens[,c("den_id", "in_park")]))

# And finally pull a shapefile of our study area
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(study_area))

##### Pull verification data #####
# Then pull our verification data
# Even though we've manually verified the den data to 
# be the best it can be, we're still going to compare
# just the autogenerated GIS forestry data for both
# groups, to ensure it's an apples-to-apples 
# comparison.
fv <- withr::with_dir(rprojroot::find_root('_targets.R'),
                            tar_read(forestry_verifications_full)) # read data verification results
pfv <- withr::with_dir(rprojroot::find_root('_targets.R'), 
                       tar_read(pseudo_forestry_verifications_full)) # pseudo data verification results

# Merge verification data to the spatial data
f_analysis <- merge(f_analysis, fv, by.x = "sample_id_forest", by.y = "sample_id") # again, merging GIS forestry data for the year PRIOR to the den status year
pseudo_dens <- merge(pseudo_dens, pfv, by = "sample_id")


##### Pull DEM data #####
# For each pseudo den, we have elevation, slope (%),
# and slope aspect, extracted from the BC CDED 30m data.
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(pseudo_dens_dem))
pseudo_dens_dem <- unique(pseudo_dens_dem)

pseudo_dens <- merge(pseudo_dens, pseudo_dens_dem)

rm(pseudo_dens_dem)

# And also clean the DEM data fields in the real data
f_analysis[["elevation_m"]][which(f_analysis$elevation_m == 999)] <- NA
f_analysis[["slope_pct"]][which(f_analysis$slope_pct > 200)] <- NA
f_analysis[["slope_aspect"]][which(f_analysis$slope_aspect > 360)] <- NA

##### Pull age class data #####
# For each pseudo den, we have the 2023 VRI age class.
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(pseudo_dens_age_class))
pseudo_dens_age_class <- sf::st_drop_geometry(pseudo_dens_age_class)

pseudo_dens <- merge(pseudo_dens, pseudo_dens_age_class, all.x = TRUE)

rm(pseudo_dens_age_class)

# And also clean up the age class field in the real data
# This will produce some NAs
pseudo_dens$proj_age_class_cd_1 <- as.numeric(pseudo_dens$proj_age_class_cd_1)
f_analysis$proj_age_class_cd_1 <- as.numeric(substring(f_analysis$age_class, 1, 1))



##### % age class #####
# already did this for f_analysis in one of the first chunks
pseudo_prct <- withr::with_dir(rprojroot::find_root('_targets.R'),
                            tar_read(pseudo_prct_age_class_1.5km))

# Bin the data into groups (sort of ~young, old, medium, very old)
pseudo_prct$lt_3 <- rowSums(pseudo_prct[,c("age_class_1", "age_class_2")])
pseudo_prct$gt_3 <- rowSums(pseudo_prct[,6:12])
pseudo_prct$three_to_7 <- rowSums(pseudo_prct[,6:10])
pseudo_prct$gt_8 <- rowSums(pseudo_prct[,c("age_class_8", "age_class_9")])

# Merge with den status - so we can look at percent of each
# age class with activity status
pseudo_dens <- merge(pseudo_dens, pseudo_prct, by = "sample_id")
pseudo_dens <- dplyr::select(pseudo_dens, -den_id.y, -year.y)
names(pseudo_dens)[grep("den_id.x", names(pseudo_dens))] <- "den_id"
names(pseudo_dens)[grep("year.x", names(pseudo_dens))] <- "year"


##### Road density #####
# `road_density` already loaded in the first chunk
withr::with_dir(rprojroot::find_root('_targets.R'), tar_load(pseudo_road_density))

pseudo_road_density <- pseudo_road_density[,c("sample_id", "road_density_m2")]

# Note the all.x = TRUE. If the den isn't in
# the road density df_analysis, the road density is 0.
pseudo_dens <- merge(pseudo_dens, pseudo_road_density, by = "sample_id", all.x = TRUE)
pseudo_dens$road_density_m2 <- ifelse(is.na(pseudo_dens$road_density_m2), 0, pseudo_dens$road_density_m2)

##### Merge datasets #####
pseudo_dens$type <- "Pseudo"
f_analysis$type <- "Real"

full <- dplyr::bind_rows(pseudo_dens, f_analysis)

# We'll use the den year sample_id (rather than the forestry year sample id)
# as the default sample id
full$sample_id <- ifelse(is.na(full$sample_id), full$sample_id_den, full$sample_id)

# Subset to only columns we care about + full observations
full <- full |> 
  dplyr::select(sample_id, den_id, type, year,
                den_status, den_status_binary,
                region, latitude, longitude, in_park,
                elevation_m, slope_pct, slope_aspect,
                proj_age_class_cd_1,
                prop_forest_60m, dist_lt40, dist_gt40, dist_road,
                age_class_1:road_density_m2)

# Filter to only 2019:2023
full <- full[which(full$year %in% c(2019:2024)), ]

#full <- na.omit(full) # this eliminates a lot of dens with incomplete DEM (slope, elevation, etc.) data

# Rename park variable so it's easier to tell what it is on plots
full$in_park <- ifelse(full$in_park, "in_park", "not_in_park")


##### Distance from edge #####
# Next, make just one 'distance from edge' variable
full$dist_from_edge <- ifelse(full$dist_lt40 > 0, full$dist_lt40, full$dist_gt40)
#hist(full$dist_from_edge)
full$log_dist_from_edge <- log(full$dist_from_edge + 1)
#hist(full$log_dist_from_edge)

# Next, say if it's in old, new, or edge of forest.
full <- full |> dplyr::mutate(dist_lt40 = round(dist_lt40, 0),
                              dist_gt40 = round(dist_gt40, 0),
                              age = dplyr::case_when((dist_lt40 == dist_gt40) ~ "on_edge",
                                               (dist_lt40 > dist_gt40) ~ "within_old",
                                               (dist_lt40 < dist_gt40) ~ "within_young"))

```


---

`r paste0(round(sum(pseudo_dens$in_park) / nrow(pseudo_dens) * 100, 0), "%")` of the randomly sampled 'pseudo dens' fall within park lands or protected areas, in contrast to only `r paste0(round(sum(dens$in_park) / nrow(dens) * 100, 0), "%")` of the dens sampled in this study.


```{r pseudo map}
##### Map #####

# Set up our map bounding box
bbox <- st_bbox(study_area)
bbox[1:2] <- bbox[1:2] - 10000 # add a 10km buffer to our bounding box
bbox[3:4] <- bbox[3:4] + 10000 # add a 10km buffer to our bounding box

# Pull BC shapefile from rnaturalearth
bc <- rnaturalearth::ne_states(country = "canada")
bc <- bc[bc$name == "British Columbia", ]
bc <- st_transform(bc, 3005) # BC Albers projection
bc <- st_crop(bc, bbox) # crop to our bounding box area

# Plot time!
# For the life of me can't figure out why the point size is not
# changing but that's a low priority thing to fix
# COMMENTED OUT TO SPEED UP RUNNING THIS SCRIPT
ggplot() +
  geom_sf(data = bc, fill = "grey") +
  geom_sf(data = study_area, fill = "darkgrey") +
  geom_sf(data = parks, fill = "lightgreen", alpha = 0.4) +
  geom_sf(data = pseudo_dens,
          aes(shape = in_park),
          stroke = NA,
          alpha = 0.3,
          size = 0.2) +
  geom_sf(data = dens,
          aes(shape = in_park),
          color = "purple",
          stroke = NA,
          size = 0.2) +
  coord_sf(xlim = c(bbox[1], bbox[3]),
           ylim = c(bbox[2], bbox[4])) +
  theme(legend.position = "none") +
  labs(caption = "Triangle shape indicates the den falls within a provincial/federal park or protected area.")
```

